Containerization of CMS Applications with Docker

### 2015-04-08 paper.tex
changes

\abstract{Clouds and virtualisation offer typical answers to the needs of large-scale computing centres to satisfy diverse sets of user communities in terms of architecture, OS, etc. On the other hand, solutions like Docker seems to emerge as a way to rely on Linux kernel capabilities to package only the applications and the development environment needed by the users, thus solving several resource management issues related to cloud-like solutions. In this paper, an exploratory work done based on the workflows of the CMS experiment at the LHC accelerator at CERN is presented. Progresses towards the deployment of a full data center via exploiting containerised software stacks for CMS will be reported and discussed.}

@@

Docker is an open platform to build, ship and run distributed applications. It is a container-based virtualisation framework that uses Linux containers (LXC), offering ease of packaging and deploying applications, a system based on lightweight containers, the use of UnionFS, as week as image storing and sharing via Docker Index/Hub. Recently, the commercial sector is showing a large set of use-cases and adoptions. Some examples are Spotify, that uses Docker for continuous delivery, service testing and deployment; Baidu, that uses Docker as a Platform-as-a-Service (PaaS), profiting of its flexibility for many framework and applications; eBay, that uses Docker for its easy application deployment, and for continuous integration process; and many more. Recently, Docker has reached 120k lines of code, about 10k commits and roughy 600 core contributors, of which about 95\% work outside Docker.

\section{Docker and CMS}

Virtualization can be achieved in several ways, depending on the desired (or acceptable) level of invasiveness. A first option is chroot or cgroups, a second options is the use of containers, and a third option is the use of virtual machines. In the latter case, virtual machines imply a high overhead in both CPU and memory, plus a non negligible time needed for creating the images, booting up and for snapshotting. Docker provides CPU, memory and filesystem isolation, and can run different processes on the same kernel but with completely different runtime (e.g. one can run SLC5 on SLC6 or on Ubuntu).

Another interesting aspect is the availability of a service like Docker Hub (hub.docker.com). It offers the management of the lifecycle of distributed applications with cloud services for building and sharing containers and automating workflows. It allows to browse, search, push, pull repositories (about 100 official images, ~45k public images), with easy mechanisms to retrieve (/upload) pre-build images via \texttt{docker pull (/push)} commands. Commands in a given image can be executed via \texttt{docker run}:
they only share the kernel with the host OS (slc6, in the example case below):
%
\begin{verbatim}
        $ docker run -it centos:centos7 cat /etc/redhat-release
        CentOS Linux release 7.0.1406 (Core)
\end{verbatim}
%
The images can be built using so called ``Dockerfiles'', e.g.:
%
\begin{verbatim}
FROM cmssw/slc6-vanilla
    RUN yum -y update && yum -y install rubygems ruby-devel \
                                      gcc ruby193
    RUN echo "gem: --no-ri --no-rdoc" > ~/.gemrc && \
        gem install puppet && \
        gem install librarian-puppet -v 1.0.9
    CMD /bin/bash
\end{verbatim}
%
All images are layered, there is no need to re-download previously downloaded layers (e.g. each of the above statements is a layer). Several CMS examples are documented~\cite{DockerGiulio}.

>>>>

\abstract{Clouds and virtualisation offer typical answers to the needs of large-scale computing centres to satisfy diverse sets of user communities in terms of architecture, OS, etc. On the other hand, solutions like Docker seems to emerge as a way to rely on Linux kernel capabilities to package only the applications and the development environment needed by the users, thus solving several resource management issues related to cloud-like solutions. In this paper, an exploratory work done based on the workflows of the CMS experiment at the LHC accelerator at CERN is presented. Progresses towards the deployment of a full data center via exploiting \emph{container-ised} software stacks for CMS will be reported and discussed.}

@@

Docker is an open platform to build, ship and run distributed applications. It is a container-based virtualisation framework that uses Linux containers (LXC) as the core technology, with a large set of tools allowing for large scale container handling, shipping and deployment. It allows for easy container contextualization, and for inter-container communications. Docker utilization is becoming the de-facto standard for container based virtualisation, and is the recommended solution endorsed by most heavy weight IT providers. Some examples are Spotify, that uses Docker for continuous delivery, service testing and deployment; Baidu, that uses Docker as a Platform-as-a-Service (PaaS), profiting of its flexibility for many framework and applications; eBay, that uses Docker for its easy application deployment, and for continuous integration process; and many more. Recently, Docker has reached 120k lines of code, about 10k commits and roughy 600 core contributors, of which about 95\% work outside Docker.

\section{Docker as a tool for light virtualization}

Virtualization can be achieved in several ways, depending on the desired (or acceptable) level of invasiveness. Setting aside solutions when the whole system hardware is emulated (like QEMU), the industry reference technologies are full system virtualization (OpernStack, OpenNebula), where each virtualised machine has its own running kernel, and kernel based process isolation, which has been a possibility in Linux since the beginning, via \emph{chroot}, and more recently via \emph{c-groups} and projects like the Linux Container (\emph{LXC}). The former technology has a broader scope, being able to utilize different operating systems on the same physical host (e.g. mixing Windows and Linux machines), but pays a price in terms of resource utilization (mainly RAM, to a smaller extent CPU reduced efficiency) and access to high performance devices; it is considered an overkill in our field, where more or less all the scirntific computation are carried out on Linux machines. Linux Containers, at the heart of the Docker technology, instead of simulating complete machines, just provede process separation and sand-boxing over a common kernel all the machines share. This is particularly adequate in an environment where system standardization has already been achieved to a large extent, for example by the adoption of WLCG-provided Middleware.
Docker provides CPU, memory and filesystem isolation, and can run different processes on the same kernel but with completely different runtime (e.g. one can run SLC5 on SLC6 or on Ubuntu).

Docker adds on top of LXC a complete series of tools for image storing, deployment, testing and contextualization, as well as for container handling (start/stop/...) and communications; derived images can be based on exiting images, with Docker taking care to save only the deltas between them. It also add a public repository (Docker Hub, =hub.docker.com=), which allows for safe image storing and easy deployment for open software projects, which provides already cooked images for most of the widespread linux based operating systems.

Currently Docker Hub hosts  ~45k publicly accessible images, and starting a new container on a linux based machine is as easy as a single line command:
%
\begin{verbatim}
        $ docker run -it centos:centos7 cat /etc/redhat-release
        CentOS Linux release 7.0.1406 (Core)
\end{verbatim}
%
The images can be built using so called ``Dockerfiles'', e.g.:
%
\begin{verbatim}
FROM cmssw/slc6-vanilla
    RUN yum -y update && yum -y install rubygems ruby-devel \
                                      gcc ruby193
    RUN echo "gem: --no-ri --no-rdoc" > ~/.gemrc && \
        gem install puppet && \
        gem install librarian-puppet -v 1.0.9
    CMD /bin/bash
\end{verbatim}
%
All images are layered, there is no need to re-download previously downloaded layers (e.g. each of the above statements is a layer). Several CMS examples are documented~\cite{DockerGiulio}.
